---
title: "Lab 1 - stringer, here"
author: "Madeline Berger"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here) #helps with project organization
library(janitor)
library(kableExtra)

```

Read in noaa_fisheries.csv data
```{r}

#us_landings <- read_csv("data/noaa_fisheries.csv") #here package helps you get around having to do this

us_landings <- read_csv(here("data","noaa_fisheries.csv"))

```

Now, make a tidy version of the data frame:
-clean_names
-str_remove
-parse_nubmer
```{r}

landings_tidy <- us_landings %>% 
  janitor::clean_names() %>% 
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_num = parse_number(dollars_usd))

```

Let's just get some information for salmon:
-str_remove
-str_detect + filter
-separate
```{r}
salmon_landings <- landings_tidy %>% 
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>% 
  filter(str_detect(afs_clean, pattern = "salmon")) %>% 
  separate(afs_clean, into = c("group","species"), sep = ",")
  
```

Find some grouped summary data:

Find annual total US landings and dollar value (summing across all states), for each type of salmon using `group_by` + `summarize`
```{r}

salmon_summary <- salmon_landings %>% 
  group_by(year, species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_num)
  )

```

Make a graph!
```{r}
salmon_landings_graph <- ggplot(data = salmon_summary,
                                aes(x = year, y = tot_landings)) +
  geom_line(aes(color = species))+
  theme_minimal()

salmon_landings_graph

```

Now we'll export our graph with `ggsave`
```{r}
ggsave(plot = salmon_landings_graph,
       here("figures","us_salmon_mb.png"),
       height = 5,
       width = 8)



```

Make a nice kable table:
```{r}

salmon_first_five <- salmon_summary %>% 
  head(5)

```

Use kable:
```{r}
kable(salmon_first_five) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```

